SIMD\slash SIMT-vectorized Sparse Chemical Kinetic Jacobian\slash Thermo-chemical Source Term Evaluation

### 2018-02-19 paper.tex
wide->shallow

%====================================================================
\begin{abstract} % not to exceed 200 words
A code generation platform for single-instruction, multiple-data (SIMD) and single-instruction, multiple thread (SIMT) vectorized thermo-chemical source term and sparse chemical kinetic Jacobian evaluation was developed and validated for a wide range of chemical kinetic models.
\todo[inline]{this}
\end{abstract}

@@

\begin{figure}[htb]
  \centering
  \begin{minipage}{0.6\linewidth}
    \begin{subfigure}[t]{\textwidth}
	\includegraphics[width=\textwidth]{row_major_split.pdf}
	\caption{Row-major, wide-vectorized data ordering}
	\label{F:row_major_split}
    \end{subfigure}
    \\
    \begin{subfigure}[t]{\textwidth}
	\includegraphics[width=\textwidth]{column_major_split.pdf}
	\caption{Column-major, deep-vectorized data ordering}
	\label{F:column_major_split}
    \end{subfigure}
  \end{minipage}
  \caption{Vectorized data ordering patterns}
  \label{F:vector_data}
\end{figure}

@@

In~\cref{F:intel_source_noorm}, the mean runtime per initial condition is shown for both the~\avx/ and~\sse/ CPUs, using both Intel OpenCL and OpenMP.
This normalization is chosen to account for the varying numbers of conditions in the PaSR databases for each model (\cref{T:error}).
For both CPUs, the OpenMP implementation is the slowest over all models; interestingly the parallel (non-vectorized) Intel OpenCL code is slightly faster than OpenMP in all cases.\todo{rerun OpenMP w/ mtune=native}
As expected, the \avx/ machine is moderately faster than the \sse/ CPU; additionally, the wide-vectorized OpenCL code is significantly faster than either the OpenMP or unvectorized OpenCL codes on both processors.

In~\cref{F:intel_source}, the extent of this speedup is detailed; we note that the speedup displayed is calculated per-machine, such that the speedup of the wide-vectorized OpenCL code can be readily assessed.
On both machines, the unvectorized OpenCL code is \textasciitilde\SIrange{1.25}{1.5}{$\times$} faster than the baseline parallel OpenMP code.
Additionally, the wide-vectorized OpenCL code is \textasciitilde\SIrange{2.5}{3}{$\times$} and \textasciitilde\SIrange{3.5}{4}{$\times$} faster than the OpenMP code for the \sse/ and \avx/ machines respectively.
In contrast,~\cref{F:pocl_source} shows the runtime of deep, wide and unvectorized OpenCL codes on the POCL runtime, as compared to OpenMP parallelization.
No speedup is achieved for either vectorization type on either CPU; indeed the OpenMP is faster in both cases.
The reasons for this lack of vectorization are quite technical and somewhat outside the scope of this work, however briefly speaking vectorization might be achieved on POCL via a combination of relaxed math (``-cl-fast-relaxed-math''), use of explicit OpenCL vector types and newer versions of LLVM (specifically, potential improvments to the vectorization module).
This will be explored more in future work, but for the moment POCL is still quite useful as a validation tool.

\begin{figure}[htbp]
   \centering
  \begin{subfigure}[t]{0.48\linewidth}
      \includegraphics[width=\textwidth]{intel_source_nonorm.pdf}
      \caption{The mean runtime per condition for each chemical model using the Intel OpenCL and OpenMP on both CPUs studied.}
      \label{F:intel_source_noorm}
  \end{subfigure}
  \hfill
  \begin{subfigure}[t]{0.48\linewidth}
      \includegraphics[width=\textwidth]{intel_source.pdf}
      \caption{The speedup achieved over the baseline OpenMP parallelization by both the unvectorized and wide-vectorized Intel OpenCL codes.}
      \label{F:intel_source}
  \end{subfigure}
  \\
  \begin{subfigure}[t]{0.48\linewidth}
      \includegraphics[width=\textwidth]{pocl_source_nonorm.pdf}
      \caption{The mean runtime per condition of the Portable OpenCL runtime compared to OpenMP parallelization.}
      \label{F:pocl_source}
  \end{subfigure}
 \caption{Mean runtimes per condition and speedups achieved by the various CPU OpenCL runtimes compared to OpenMP parallelization for each chemical model studied. The names in the legends correspond to the identifiers listed in~\cref{t:cpus}}
 \label{F:cpu_source}
\end{figure}

@@

\begin{figure}[htbp]
   \centering
  \begin{subfigure}[t]{0.48\linewidth}
      \includegraphics[width=\textwidth]{gpu_source_scaling.pdf}
      \caption{The mean runtime per condition for each chemical model on the \gpunew/ GPU as a function of the number of initial conditions tested.}
      \label{F:gpu_source_scaling}
  \end{subfigure}
  \hfill
  \begin{subfigure}[t]{0.48\linewidth}
      \includegraphics[width=\textwidth]{gpu_source_speedup.pdf}
      \caption{The speedup achieved on the \gpunew/ versus the \gpuold/ GPU; the names correspond to the identifiers listed in~\cref{t:gpus}}
      \label{F:gpu_source_speedup}
  \end{subfigure}
  \caption{\texttt{pyJac} source term evaluation performance on the NVIDIA GPUs}
  \label{F:gpu_source}
\end{figure}

@@

In contrast,~\Cref{F:source_cvsf} shows large speedups of ``C''-ordered data over ``F''-ordered data on the \avx/ machine; the speedup presented is calculated per language, i.e., the roughly \SIrange{1.5}{2}{$\times$} speedup of the ``C''-ordered OpenMP implementation is relative to the ``F''-order OpenMP baseline while the OpenCL speedups are given with respect to the ``F''-ordered unvectorized OpenCL code.
Additionally, we note that the ``C''-ordered wide-vectorization in~\cref{F:source_cvsf} utilizes the vectorized-data ordering described in~\cref{S:data}.
In general, the data-ordering has a much more profound affect on the vectorized-OpenCL codes; this is due to the nature of SIMD-operations, which are much more efficient when operating on contiguous (in memory) data rather than incurring an expensive memory ``gather'' operation.
The ``C''-ordered unvectorized OpenCL code (which uses the default unvectorized row-major data ordering) is \textasciitilde\SIrange{1.5}{2.5}{$\times$} faster than the unvectorized ``F''-ordered code, and is in some cases almost as fast as the ``F''-ordered wide-vectorized OpenCL implementation.
However, the ``C''-ordered wide-vectorization achieves speedups of \textasciitilde\SIrange{4}{7}{$\times$} demonstrating the value of the vectorized data ordering for CPU execution.

\Cref{F:source_gpu_cvsf} shows the effect of the ``C'' versus ``F''-ordering on source term evaluation performance on both GPUs.
The speedup reported is calculated per-GPU, such that the effect of data ordering can be seen on each individual GPU.
In general, the ``C''-ordered wide-vectorization is \textasciitilde\SIrange{1.05}{1.2}{$\times$} faster on the \gpunew/ GPU while the ``C'' and ``F''-ordered data are nearly equivalent for the \gpuold/ GPU, except for the isopentanol model where the ``F''-ordered data is \textasciitilde\SI{1.4}{$\times$} faster (though, we note that there was high variablity in this case). \todo{again, pending rerun on \gpuold/}
The ``C''-ordered speedup on the \gpunew/ GPU is counter to what one might expect, as typically speaking coalesced memory access in a wide-vectorization is easier to achieve if ``F''-ordering is used (see~\cref{S:data}).
However, the vectorized-data ordering here ensures that not only are memory accesses aligned to the vector width size (and thus, likely coalesced) but the improved data-locality; coupled with the larger cache sizes of the \gpunew/ GPU result in this modest speedup for ``C''-data ordering.

\begin{figure}[htbp]
   \centering
  \begin{subfigure}[t]{0.48\linewidth}
      \includegraphics[width=\textwidth]{source_conpvsconv.pdf}
      \caption{The mean runtime per condition for each chemical model using both the \conp/ and \conv/ formulations on Intel OpenCL and OpenMP for \avx/.}
      \label{F:source_conpvsconv}
  \end{subfigure}
  \hfill
  \begin{subfigure}[t]{0.48\linewidth}
      \includegraphics[width=\textwidth]{source_cvsf.pdf}
      \caption{The speedup achieved by ``C'' ordering over ``F'' ordering for Intel OpenCL and OpenMP on the \avx/ CPU.  The speedup presented is calculated per-language (OpenMP and OpenCL) to better assess the effect of the data-ordering.}
      \label{F:source_cvsf}
  \end{subfigure}
  \\
  \begin{subfigure}[t]{0.48\linewidth}
      \includegraphics[width=\textwidth]{source_vector_width.pdf}
      \caption{The effect of vector-width on ``C''-ordered wide-vectorized Intel OpenCL source term evaluations in the \avx/ CPU.}
      \label{F:source_vector_width}
  \end{subfigure}
  \hfill
  \begin{subfigure}[t]{0.48\linewidth}
      \includegraphics[width=\textwidth]{source_gpu_cvsf.pdf}
      \caption{The affect on performance of ``C'' vs ``F''-ordering for the wide-vectorized NVIDIA OpenCL code on both GPUs with a vector width of 128.  The speedup presented is calculated per-GPU to better assess the effect of the data-ordering on each GPU.}
      \label{F:source_gpu_cvsf}
  \end{subfigure}
  \caption{The effect of the \conp/ and \conv/ formulations, as well as ``C'' and ``F'' data-ordering on source term evaluation performance in \texttt{pyJac}.
	   It is noted the wide-vectorized ``C'' ordered OpenCL case corresponds to the vectorized data ordering described in~\cref{S:data}}
  \label{f:source_permutate}
\end{figure}

@@

\Cref{F:source_parallel_scaling} shows the strong parallel scaling efficiency of source rate evaluation in \texttt{pyJac} on the \avx/ machine for both the wide-vectorized Intel OpenCL and OpenMP codes.
In general, the \ce{H2}\slash\ce{CO} mechanism has the worst scaling efficiency for Intel OpenCL; although the mechanism scalings vary somewhat for OpenMP, there is no clear trend.
It is likely the poor scaling of the \ce{H2}\slash\ce{CO} for Intel OpenCL results from both its relatively small size, and few falloff\slash chemically activated reaction (and in particular, the additional expensive logarithm and exponential evaluations that accompany them); i.e., there is simply not enough work to fully utilize the SIMD intructions.
Additionally, OpenMP tends to have higher scalings then the wide-vectorized OpenCL code, e.g., \textasciitilde\num{0.8} and \textasciitilde\num{0.7} for \num{4} and \num{8} CPU cores respectively, compared to just \textasciitilde\num{0.7} and \textasciitilde\num{0.5} for OpenCL; although not pictured (to keep the figure readable), the unvectorized Intel OpenCL code has only slightly worse scaling than the OpenMP code, hence the SIMD-vectorization itself the cause of this poorer scaling.
This is again likely due to relatively small amounts of work associated with source rate evaluation; the longest runtime (the isopentanol model on a single core) takes only \textasciitilde$\SI{2}{\sec}$ for OpenCL, and \textasciitilde$\SI{9}{\sec}$ for OpenMP.
By the time 4--8 CPU cores are utilized, the runtime of wide-vectorized OpenCL code can be measured in hundreds of milliseconds, and parallel overhead begins to dominate, decreasing efficiency.
In contrast, the OpenMP code requires \textasciitilde\num{4} cores to run as fast as the wide-vectorized OpenCL code in this case, hence parallel overhead concerns become increasingly important for 8--16 cores.
Indeed, we will see that the strong scalings for sparse Jacobian evaluation---the most computationally intensive task in this work for which the efficiency will investigated---are far more similar between Intel OpenCL and OpenMP.

The SIMD efficiency is defined as:
\begin{equation}
 \label{e:simd_efficiency}
 \varepsilon_{\text{SIMD}} = \frac{\bar{t}_{\text{unvec}}}{W \bar{t}_{\text{wide}}}
\end{equation}
where $\bar{t}_{\text{unvec}}$ is the mean runtime per condition of the unvectorized OpenCL code, $\bar{t}_{\text{wide}}$ the same for the wide-vectorized OpenCL code, and $W$, the vector width reported in number of double operations in~\cref{t:cpus}
This measure compares the actual speedup due to wide-vectorization compared to the nominal vector-width of the machine in question.
\Cref{F:source_simd_scaling} shows the SIMD efficiency of source term evaluation in \texttt{pyJac} on a single core of the \avx/ machine; the larger models have higher SIMD-efficiency, \textasciitilde\num{0.75}, and the smaller models have lower SIMD-efficiency, \textasciitilde\num{0.625}.
This again demonstrates that the SIMD-vectorization becomes efficient as the amount of work to perform increases (with increasing model size).
Interestingly,~\cref{F:source_sse_simd_scaling} shows the SIMD efficiency on the \sse/ machine as greater than one; this is likely caused by a combination of use of an OpenCL vector-width greater than the native CPU vector-width (i.e., \num{8} v.s., \num{2}), and the improved data-locality for the vectorized-data ordering, as discussed in~\cref{S:data}, and results a modest \textasciitilde\SIrange{10}{25}{\percent} improvement over the nominal vector width.

@@

\section{Conclusions}
In this work, automatically generated OpenCL codes for shallow SIMD-vectorized thermo-chemical source term evaluation were developed and validated against Cantera~\cite{Cantera} for a wide range of chemical kinetic models~\cite{Burke:2011fh,smith_gri-mech_30,Wang:2007,Sarathy:2013jr}.
Significant speedups of up to \SIrange{1.90}{2.45}{$\times$} over a baseline SIMT-vectorized code were observed.
Two data-ordering schemes were investigated, showing a clear performance benefit for use of the ``C'' (row-major) ordering for the shallow vectorized code.
Further, this study suggests that a deep SIMD-vectorized code---currently under development---may see even greater accelerations due to expected increases in cache-hit rates.
Larger vector-widths were found to provide accelerations for the smaller models studied, but the speedup was similar over all vector-widths tested for the largest models investigated.
Finally, the strong parallel scaling efficiency was examined for the shallow vectorized code; the larger models exhibited higher scaling efficiency, possibly due to failure of the \ce{H2}\slash\ce{CO} model to saturate the CPU throughput.
Future extensions of this work will include development of a SIMD\slash SIMT-accelerated sparse-analytical Jacobian code targeted at CPUs, GPUs, and MICs.

>>>>

%====================================================================
\begin{abstract} % not to exceed 200 words
A code generation platform for single-instruction, multiple-data (SIMD) and single-instruction, multiple thread (SIMT) vectorized thermo-chemical source term and sparse chemical kinetic Jacobian evaluation was developed and validated for a shallow range of chemical kinetic models.
\todo[inline]{this}
\end{abstract}

@@

\begin{figure}[htb]
  \centering
  \begin{minipage}{0.6\linewidth}
    \begin{subfigure}[t]{\textwidth}
	\includegraphics[width=\textwidth]{row_major_split.pdf}
	\caption{Row-major, shallow-vectorized data ordering}
	\label{F:row_major_split}
    \end{subfigure}
    \\
    \begin{subfigure}[t]{\textwidth}
	\includegraphics[width=\textwidth]{column_major_split.pdf}
	\caption{Column-major, deep-vectorized data ordering}
	\label{F:column_major_split}
    \end{subfigure}
  \end{minipage}
  \caption{Vectorized data ordering patterns}
  \label{F:vector_data}
\end{figure}

@@

In~\cref{F:intel_source_noorm}, the mean runtime per initial condition is shown for both the~\avx/ and~\sse/ CPUs, using both Intel OpenCL and OpenMP.
This normalization is chosen to account for the varying numbers of conditions in the PaSR databases for each model (\cref{T:error}).
For both CPUs, the OpenMP implementation is the slowest over all models; interestingly the parallel (non-vectorized) Intel OpenCL code is slightly faster than OpenMP in all cases.\todo{rerun OpenMP w/ mtune=native}
As expected, the \avx/ machine is moderately faster than the \sse/ CPU; additionally, the shallow-vectorized OpenCL code is significantly faster than either the OpenMP or unvectorized OpenCL codes on both processors.

In~\cref{F:intel_source}, the extent of this speedup is detailed; we note that the speedup displayed is calculated per-machine, such that the speedup of the shallow-vectorized OpenCL code can be readily assessed.
On both machines, the unvectorized OpenCL code is \textasciitilde\SIrange{1.25}{1.5}{$\times$} faster than the baseline parallel OpenMP code.
Additionally, the shallow-vectorized OpenCL code is \textasciitilde\SIrange{2.5}{3}{$\times$} and \textasciitilde\SIrange{3.5}{4}{$\times$} faster than the OpenMP code for the \sse/ and \avx/ machines respectively.
In contrast,~\cref{F:pocl_source} shows the runtime of deep, shallow and unvectorized OpenCL codes on the POCL runtime, as compared to OpenMP parallelization.
No speedup is achieved for either vectorization type on either CPU; indeed the OpenMP is faster in both cases.
The reasons for this lack of vectorization are quite technical and somewhat outside the scope of this work, however briefly speaking vectorization might be achieved on POCL via a combination of relaxed math (``-cl-fast-relaxed-math''), use of explicit OpenCL vector types and newer versions of LLVM (specifically, potential improvments to the vectorization module).
This will be explored more in future work, but for the moment POCL is still quite useful as a validation tool.

\begin{figure}[htbp!]
   \centering
  \begin{subfigure}[t]{0.48\linewidth}
      \includegraphics[width=\textwidth]{intel_source_nonorm.pdf}
      \caption{The mean runtime per condition for each chemical model using the Intel OpenCL and OpenMP on both CPUs studied.}
      \label{F:intel_source_noorm}
  \end{subfigure}
  \hfill
  \begin{subfigure}[t]{0.48\linewidth}
      \includegraphics[width=\textwidth]{intel_source.pdf}
      \caption{The speedup achieved over the baseline OpenMP parallelization by both the unvectorized and shallow-vectorized Intel OpenCL codes; the speedup is presented on per-machine basis, e.g., the \avx2/ shallow-vectorized code speedup is relative to OpenMP on the same CPU.}
      \label{F:intel_source}
  \end{subfigure}
  \\
  \begin{subfigure}[t]{0.48\linewidth}
      \includegraphics[width=\textwidth]{pocl_source_nonorm.pdf}
      \caption{The mean runtime per condition of the Portable OpenCL runtime compared to OpenMP parallelization.}
      \label{F:pocl_source}
  \end{subfigure}
 \caption{Mean runtimes per condition and speedups achieved by the various CPU OpenCL runtimes compared to OpenMP parallelization for each chemical model studied. The names in the legends correspond to the identifiers listed in~\cref{t:cpus}.}
 \label{F:cpu_source}
\end{figure}

@@

\begin{figure}[htbp!]
   \centering
  \begin{subfigure}[t]{0.48\linewidth}
      \includegraphics[width=\textwidth]{gpu_source_scaling.pdf}
      \caption{The mean runtime per condition for each chemical model on the \gpunew/ GPU as a function of the number of initial conditions tested.}
      \label{F:gpu_source_scaling}
  \end{subfigure}
  \hfill
  \begin{subfigure}[t]{0.48\linewidth}
      \includegraphics[width=\textwidth]{gpu_source_speedup.pdf}
      \caption{The speedup achieved on the \gpunew/ versus the \gpuold/ GPU; the names correspond to the identifiers listed in~\cref{t:gpus}}
      \label{F:gpu_source_speedup}
  \end{subfigure}
  \caption{\texttt{pyJac} source term evaluation performance on the NVIDIA GPUs}
  \label{F:gpu_source}
\end{figure}

@@

In contrast,~\Cref{F:source_cvsf} shows large speedups of ``C''-ordered data over ``F''-ordered data on the \avx/ machine; the speedup presented is calculated per language, i.e., the roughly \SIrange{1.5}{2}{$\times$} speedup of the ``C''-ordered OpenMP implementation is relative to the ``F''-order OpenMP baseline while the OpenCL speedups are given with respect to the ``F''-ordered unvectorized OpenCL code.
Additionally, we note that the ``C''-ordered shallow-vectorization in~\cref{F:source_cvsf} utilizes the vectorized-data ordering described in~\cref{S:data}.
In general, the data-ordering has a much more profound affect on the vectorized-OpenCL codes; this is due to the nature of SIMD-operations, which are much more efficient when operating on contiguous (in memory) data rather than incurring an expensive memory ``gather'' operation.
The ``C''-ordered unvectorized OpenCL code (which uses the default unvectorized row-major data ordering) is \textasciitilde\SIrange{1.5}{2.5}{$\times$} faster than the unvectorized ``F''-ordered code, and is in some cases almost as fast as the ``F''-ordered shallow-vectorized OpenCL implementation.
However, the ``C''-ordered shallow-vectorization achieves speedups of \textasciitilde\SIrange{4}{7}{$\times$} demonstrating the value of the vectorized data ordering for CPU execution.

\Cref{F:source_gpu_cvsf} shows the effect of the ``C'' versus ``F''-ordering on source term evaluation performance on both GPUs.
The speedup reported is calculated per-GPU, such that the effect of data ordering can be seen on each individual GPU.
In general, the ``C''-ordered shallow-vectorization is \textasciitilde\SIrange{1.05}{1.2}{$\times$} faster on the \gpunew/ GPU while the ``C'' and ``F''-ordered data are nearly equivalent for the \gpuold/ GPU, except for the isopentanol model where the ``F''-ordered data is \textasciitilde\SI{1.4}{$\times$} faster (though, we note that there was high variablity in this case). \todo{again, pending rerun on \gpuold/}
The ``C''-ordered speedup on the \gpunew/ GPU is counter to what one might expect, as typically speaking coalesced memory access in a shallow-vectorization is easier to achieve if ``F''-ordering is used (see~\cref{S:data}).
However, the vectorized-data ordering here ensures that not only are memory accesses aligned to the vector width size (and thus, likely coalesced) but the improved data-locality; coupled with the larger cache sizes of the \gpunew/ GPU result in this modest speedup for ``C''-data ordering.

\begin{figure}[htbp]
   \centering
  \begin{subfigure}[t]{0.48\linewidth}
      \includegraphics[width=\textwidth]{source_conpvsconv.pdf}
      \caption{The mean runtime per condition for each chemical model using both the \conp/ and \conv/ formulations on Intel OpenCL and OpenMP for \avx/.}
      \label{F:source_conpvsconv}
  \end{subfigure}
  \hfill
  \begin{subfigure}[t]{0.48\linewidth}
      \includegraphics[width=\textwidth]{source_cvsf.pdf}
      \caption{The speedup achieved by ``C'' ordering over ``F'' ordering for Intel OpenCL and OpenMP on the \avx/ CPU.  The speedup presented is calculated per-language (OpenMP and OpenCL) to better assess the effect of the data-ordering.}
      \label{F:source_cvsf}
  \end{subfigure}
  \\
  \begin{subfigure}[t]{0.48\linewidth}
      \includegraphics[width=\textwidth]{source_vector_width.pdf}
      \caption{The effect of vector-width on ``C''-ordered shallow-vectorized Intel OpenCL source term evaluations in the \avx/ CPU.}
      \label{F:source_vector_width}
  \end{subfigure}
  \hfill
  \begin{subfigure}[t]{0.48\linewidth}
      \includegraphics[width=\textwidth]{source_gpu_cvsf.pdf}
      \caption{The affect on performance of ``C'' vs ``F''-ordering for the shallow-vectorized NVIDIA OpenCL code on both GPUs with a vector width of 128.  The speedup presented is calculated per-GPU to better assess the effect of the data-ordering on each GPU.}
      \label{F:source_gpu_cvsf}
  \end{subfigure}
  \caption{The effect of the \conp/ and \conv/ formulations, as well as ``C'' and ``F'' data-ordering on source term evaluation performance in \texttt{pyJac}.
	   It is noted the shallow-vectorized ``C'' ordered OpenCL case corresponds to the vectorized data ordering described in~\cref{S:data}}
  \label{f:source_permutate}
\end{figure}

@@

\Cref{F:source_parallel_scaling} shows the strong parallel scaling efficiency of source rate evaluation in \texttt{pyJac} on the \avx/ machine for both the shallow-vectorized Intel OpenCL and OpenMP codes.
In general, the \ce{H2}\slash\ce{CO} mechanism has the worst scaling efficiency for Intel OpenCL; although the mechanism scalings vary somewhat for OpenMP, there is no clear trend.
It is likely the poor scaling of the \ce{H2}\slash\ce{CO} for Intel OpenCL results from both its relatively small size, and few falloff\slash chemically activated reaction (and in particular, the additional expensive logarithm and exponential evaluations that accompany them); i.e., there is simply not enough work to fully utilize the SIMD intructions.
Additionally, OpenMP tends to have higher scalings then the shallow-vectorized OpenCL code, e.g., \textasciitilde\num{0.8} and \textasciitilde\num{0.7} for \num{4} and \num{8} CPU cores respectively, compared to just \textasciitilde\num{0.7} and \textasciitilde\num{0.5} for OpenCL; although not pictured (to keep the figure readable), the unvectorized Intel OpenCL code has only slightly worse scaling than the OpenMP code, hence the SIMD-vectorization itself the cause of this poorer scaling.
This is again likely due to relatively small amounts of work associated with source rate evaluation; the longest runtime (the isopentanol model on a single core) takes only \textasciitilde$\SI{2}{\sec}$ for OpenCL, and \textasciitilde$\SI{9}{\sec}$ for OpenMP.
By the time 4--8 CPU cores are utilized, the runtime of shallow-vectorized OpenCL code can be measured in hundreds of milliseconds, and parallel overhead begins to dominate, decreasing efficiency.
In contrast, the OpenMP code requires \textasciitilde\num{4} cores to run as fast as the shallow-vectorized OpenCL code in this case, hence parallel overhead concerns become increasingly important for 8--16 cores.
Indeed, we will see that the strong scalings for sparse Jacobian evaluation---the most computationally intensive task in this work for which the efficiency will investigated---are far more similar between Intel OpenCL and OpenMP.

The SIMD efficiency is defined as:
\begin{equation}
 \label{e:simd_efficiency}
 \varepsilon_{\text{SIMD}} = \frac{\bar{t}_{\text{unvec}}}{W \bar{t}_{\text{shallow}}}
\end{equation}
where $\bar{t}_{\text{unvec}}$ is the mean runtime per condition of the unvectorized OpenCL code, $\bar{t}_{\text{shallow}}$ the same for the shallow-vectorized OpenCL code, and $W$, the vector width reported in number of double operations in~\cref{t:cpus}
This measure compares the actual speedup due to shallow-vectorization compared to the nominal vector-width of the machine in question.
\Cref{F:source_simd_scaling} shows the SIMD efficiency of source term evaluation in \texttt{pyJac} on a single core of the \avx/ machine; the larger models have higher SIMD-efficiency, \textasciitilde\num{0.75}, and the smaller models have lower SIMD-efficiency, \textasciitilde\num{0.625}.
This again demonstrates that the SIMD-vectorization becomes efficient as the amount of work to perform increases (with increasing model size).
Interestingly,~\cref{F:source_sse_simd_scaling} shows the SIMD efficiency on the \sse/ machine as greater than one; this is likely caused by a combination of use of an OpenCL vector-width greater than the native CPU vector-width (i.e., \num{8} v.s., \num{2}), and the improved data-locality for the vectorized-data ordering, as discussed in~\cref{S:data}, and results a modest \textasciitilde\SIrange{10}{25}{\percent} improvement over the nominal vector width.

@@

\section{Conclusions}
In this work, automatically generated OpenCL codes for shallow SIMD-vectorized thermo-chemical source term evaluation were developed and validated against Cantera~\cite{Cantera} for a shallow range of chemical kinetic models~\cite{Burke:2011fh,smith_gri-mech_30,Wang:2007,Sarathy:2013jr}.
Significant speedups of up to \SIrange{1.90}{2.45}{$\times$} over a baseline SIMT-vectorized code were observed.
Two data-ordering schemes were investigated, showing a clear performance benefit for use of the ``C'' (row-major) ordering for the shallow vectorized code.
Further, this study suggests that a deep SIMD-vectorized code---currently under development---may see even greater accelerations due to expected increases in cache-hit rates.
Larger vector-widths were found to provide accelerations for the smaller models studied, but the speedup was similar over all vector-widths tested for the largest models investigated.
Finally, the strong parallel scaling efficiency was examined for the shallow vectorized code; the larger models exhibited higher scaling efficiency, possibly due to failure of the \ce{H2}\slash\ce{CO} model to saturate the CPU throughput.
Future extensions of this work will include development of a SIMD\slash SIMT-accelerated sparse-analytical Jacobian code targeted at CPUs, GPUs, and MICs.
